---
title: "p8105_hw2_sc4636"
output: github_document
---

# Data Wrangling 1

```{r setup, include=FALSE}
library(tidyverse)
library(readxl)

knitr::opts_chunk$set(
  echo = TRUE, 
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_bw() + theme(legend.position = "bottom"))
```

## _Problem 1_

* Read and clean the Mr. Trash Wheel sheet
```{r, message = FALSE}
mr_trash = read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
                      range = "Mr. Trash Wheel!A2:N408") %>%
  janitor::clean_names() %>%
  rename(weight = weight_tons, volume = volume_cubic_yards) %>% 
  drop_na(dumpster) %>%
  mutate(sports_balls = as.integer(round(sports_balls)))
```  

* Read and clean precipitation data for 2017 and 2018

```{r, message = FALSE}
prcp_2017 = read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
                       range = "2017 Precipitation!A2:B14") %>%
  janitor::clean_names() %>% 
  drop_na(total) %>% 
  mutate(year = 2017)

prcp_2018 = read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
                       range = "2018 Precipitation!A2:B14") %>%
  janitor::clean_names() %>% 
  drop_na(total) %>% 
  mutate(year = 2018)
```

* Combine the precipitation datasets

```{r, message = FALSE}
prcp_1817 = bind_rows(prcp_2017, prcp_2018) %>% 
  mutate(month = month.name[month]) %>% 
  arrange(month, year, total)
```

The tidied Mr. Trash Wheel dataset has `r nrow(mr_trash)` observations of `r ncol(mr_trash)` variables, and includes information on the dumpster number, date of collection, amount of total litter and litter type, with key variables such as _`r mr_trash %>% select(dumpster:volume) %>% colnames`_.

The combined dataset for precipitation in 2017 and 2018 has `r nrow(prcp_1817)` observations of `r ncol(prcp_1817)` variables including _`r colnames(prcp_1817)`_. The total precipitation in 2018 was `r prcp_1817 %>% filter(year == 2018) %>% select(total) %>% colSums()`. The median number of sports balls in a dumpster in 2017 was `r mr_trash %>% filter(year == 2017) %>% pull(sports_balls) %>% median()`.

## _Problem 2_

* Clean the data in pols-month.csv

```{r, message = FALSE}
pols = read_csv("./data/pols-month.csv") %>% 
  janitor::clean_names() %>% 
  separate(mon, c("year", "month", "day"), "-") %>% 
  mutate_if(is.character, as.numeric) %>% 
  mutate(month = month.abb[month]) %>% 
  mutate(president = if_else(prez_dem == 1, "dem", "gop")) %>% 
  select(-prez_gop, -prez_dem, -day)
```

* Clean the data in snp.csv

```{r, message = FALSE}
snp = read_csv("./data/snp.csv") %>% 
  janitor::clean_names() %>% 
  separate(date, c("month", "day", "year"), "/") %>% 
  mutate_if(is.character, as.numeric) %>% 
  mutate(month = month.abb[month]) %>% 
  select(year, month, -day, close)
```

* Clean the data in unemployment

```{r, message = FALSE}
unemploy = read_csv("./data/unemployment.csv") %>% 
  pivot_longer(Jan:Dec, names_to = "month", values_to = "unemploy_pct") %>% 
  janitor::clean_names()
```  

* Join the datasets

```{r, message = FALSE}
pols_snp = left_join(pols, snp, by = c("year", "month"))

pols_snp_unemploy = left_join(pols_snp, unemploy, by = c("year", "month"))
```

The dataset `pols` contains `r nrow(pols)` observations of `r ncol(pols)` variables and includes information on the number of politicians at any given time with key variables _`r pols %>% select(year, month, president) %>% colnames`_. The year range for this dataset is `r pols %>% pull(year) %>% range()`.

The dataset `snp` contains `r nrow(snp)` observations of `r ncol(snp)` variables with information of the closing values on given dates. It has variables _`r colnames(snp)`_. The year range for this dataset is `r snp %>% pull(year) %>% range()`.

The dataset `unemployment` contains `r nrow(unemploy)` observations of `r ncol(unemploy)` variables indicating percentage of unemployment at different time with variables _`r colnames(unemploy)`_. The year range for this dataset is `r unemploy %>% pull(year) %>% range()`.

The joined dataset contains `r nrow(pols_snp_unemploy)` observations of `r ncol(pols_snp_unemploy)` variables. It has key variables _`r pols_snp_unemploy %>% select(-gov_gop:-rep_dem) %>% colnames`_. The year range for this dataset is `r pols_snp_unemploy %>% pull(year) %>% range()`, the same as the dataset `pols`.

## _Problem 3_

* Load and tidy the data

```{r, message = FALSE}
baby_name = read_csv("./data/Popular_Baby_Names.csv") %>%
  janitor::clean_names() %>% 
  mutate_if(is.character, str_to_title) %>% 
  mutate(ethnicity = recode(ethnicity,
                            "Asian And Paci" = "Asian And Pacific Islander",
                            "Black Non Hisp" = "Black Non Hispanic",
                            "White Non Hisp" = "White Non Hispanic")) %>% 
  distinct()
```

* Tables

```{r}
olivia_pop = baby_name %>% 
  filter(childs_first_name == "Olivia") %>% 
  select(ethnicity, year_of_birth, rank) %>%
  pivot_wider(names_from = "year_of_birth", values_from = "rank")
```

```{r, echo = FALSE}
olivia_pop %>% 
  knitr::kable(
    caption = "Table 1: Popularity of 'Olivia'",
    col.names = c("Ethnicity", 2016:2011)
  )
```

```{r}
male_pop = baby_name %>% 
  filter(gender == "Male", rank == "1") %>% 
  select(ethnicity, year_of_birth, childs_first_name) %>% 
  pivot_wider(names_from = "year_of_birth", values_from = "childs_first_name")
```

```{r, echo = FALSE}
male_pop %>% 
  knitr::kable(
    caption = "Table 2: The most popular name for male children",
    col.names = c("Ethnicity", 2016:2011)
  )
```

* Scatter plot

```{r, echo = FALSE}
male_wnp_16 = baby_name %>% 
  filter(
    gender == "Male",
    ethnicity == "White Non Hispanic",
    year_of_birth == "2016") %>% 
  ggplot(aes(x = rank, y = count, color = childs_first_name)) +
  geom_point(alpha = .8) +
  theme(legend.position = "none") +
  labs(
    title = "Popularity of baby names",
    x = "Rank of popularity", 
    y = "The number of children",
    caption = "Male, white non-hispanic children born in 2016"
  ) +
  viridis::scale_color_viridis(discrete = TRUE)

male_wnp_16
```



